\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2014-05-19 paper.tex
Rephrase, integrate HATP section, based on Rachid's contribution

Our execution controllers rely on symbolic task planning to convert long-term
desires into a succession of atomic actions. We use in our architecture the HATP
planner (for Human Aware Task Planner~\cite{Alili2008,
Alili2009,Lallement2014}).

HATP (Hierarchical Agent- based Task Planner) planning framework extends the
traditional HTN (Hierarchical Task Network) planning domain representation and
semantics by making them more suitable to elaborate plans which involve humans
and robots acting together toward a joint goal. HATP is used by the robot to
elaborate human-robot \emph{shared plans} which are then used to anticipate
human action, to propose to human to act or even to ask help from the human.

@@

An important feature of HATP is that it treats agents as ``first-class
entities'' in the domain representation language. It can therefore
distinguish between the different agents in the domain as well as
between agents and the other entities such as tables and chairs. This
facilitates a post-processing step in HATP that splits the final
solution (sequence of actions) into two (or more if there are several
humans) synchronised solution
streams, one for the robot and one for the human, so that the streams may be executed in
parallel and synchronised when necessary.

@@

The planner also generates synchronisation points between the agents.
Figure~\ref{plan_hatp1} illustrates such a plan with two streams.  Here, the
robot needs to wait for the success of the human agent's {\tt PUT} action
(estimated through the fulfilment of the action post-conditions, typically
$\neg$\stmt{GREY\_TAPE isOn TABLE}) to be able to start its own action.

@@

\caption{A plan produced by HATP with two action streams (human actions at the
top, robot actions at the bottom). The arrow between the two streams represents
a synchronization point ({\sf PUTRV} stands for {\it Put it so it is both
reachable and visible}).}

@@

To address some of those challenges HATP includes mechanisms called
"social tules" to filter plans so as to keep only those suitable for
HRI. To this end, HATP allows the specification of the following
filtering criteria:

\emph{Wasted time}: Avoids plans where the human mentioned in a plan spends a lot
of its time waiting.

\emph{Effort balancing:} Avoids plans where efforts are
not fairly distributed among the agents mentioned in a plan.  It is
indeed sometimes beneficial to balance efforts between the human and
the robot.

\emph{Control of intricacy:} Avoids plans with too many interdependencies
between the actions of agents mentioned in the plan, as a problem with
executing just one of those actions could invalidate the entire
plan. Also intricate human-robot activity may cause discomfort since
the human will find himself repeatedly in a situation where his is
waiting for the robot to act

\emph{Undesirable sequences:} Avoids plans that violate specific user-defined
sequences (for instance sequences which can be misinterpreted by the human).

Combining some of the above criteria could help yield the
following interesting \emph{shared plans}: (1) the human ends up doing a
number of the tasks, but yet the overall effort is significantly lower than what the robot puts
to do a lower number of effort-intensive tasks; (2) avoiding, when
possible, having the human wait for the robot several times, which
essentially prevents the streams from having too many causal links
between them.

 In the current implementation, the so-called social rules are
implemented by looking through all the plans produced and filtering
out the ones that do not meet the requirements specified. In the
future we intend to study algorithms that do such filtering online,
rather than after primitive solutions are found.

@@

Figure~\ref{plan_hatp2} illustrates an alternative plan to the previous 
one (Figure~\ref{plan_hatp1}) where the \emph{no wasted time} social rule is
applied. The resulting shared plan is a better plan according to a global evaluation of
the costs and rules.

>>>>

Our execution controllers rely on symbolic task planning to convert long-term
desires into a succession of atomic actions. We use in our architecture the HATP
planner (\emph{Human Aware Task Planner})~\cite{Alili2008,
Alili2009,Lallement2014}.

The HATP planning framework extends the
traditional HTN (Hierarchical Task Network) planning domain representation and
semantics by making them more suitable to elaborate plans which involve humans
and robots acting together toward a joint goal. HATP is used by the robot to
elaborate human-robot \emph{shared plans} which are then used to anticipate
human action, to propose to human to act or even to ask help from the human.

@@

An important feature of HATP is that it treats agents as ``first-class
entities'' in the domain representation language. It can therefore
distinguish between the different agents in the domain as well as
between agents and the other entities such as tables and chairs. This
facilitates a post-processing step in HATP that splits the final
solution (sequence of actions) into two (or more if there are several
humans) synchronised solution
streams, one for the robot and one for the human, so that the streams may be executed in
parallel and synchronised when necessary (Figure~\ref{plan_hatp1}).

@@

The planner also generates synchronisation points between the agents.  In the
plan depicted on Figure~\ref{plan_hatp1}, the robot needs to wait for the
success of the human agent's {\tt PUT} action (estimated through the fulfilment
of the action post-conditions, typically $\neg$\stmt{GREY\_TAPE isOn TABLE}) to
be able to start its own action.

@@

\caption{A plan produced by HATP with two action streams (human actions at the
top, robot actions at the bottom, {\sf PUTRV} stands here for {\it Put it so it is both
reachable and visible}). The arrow between the two streams represents
a synchronization point.}

@@

HATP includes mechanisms called \emph{social rules} to filter plans so as to
keep only those considered as suitable for human-robot interaction. The planner
allows the specification of the following filtering criteria:

\emph{Wasted time}: avoids plans where the human spends a lot
of its time being idle,

\emph{Effort balancing}: avoids plans where efforts are
not fairly distributed among the agents taking part to the plan.  It is
indeed sometimes beneficial to balance efforts between the human and
the robot,

\emph{Control of intricacy}: avoids plans with too many interdependencies
between the actions of agents mentioned in the plan, as a problem with
executing just one of those actions could invalidate the entire
plan. Also intricate human-robot activity may cause discomfort since
the human will find himself repeatedly in a situation where his is
waiting for the robot to act,

\emph{Undesirable sequences}: avoids plans that violate specific user-defined
sequences (for instance sequences which can be misinterpreted by the human).

Combining the above criteria, we typically yield shared plans with desirable
interaction features like the human still being engaged in a number of tasks
while its overall level of required efforts remains low, or avoiding having the
human to wait for the robot (by essentially preventing the action streams from
having too many causal links between them).

Figure~\ref{plan_hatp2} illustrates such a socially-optimised plan where the
\emph{no wasted time} social rule is applied. The resulting shared plan is
considered better than the one depicted on Figure~\ref{plan_hatp1} according to
a global evaluation of the costs and rules.

In the current implementation, the social rules are effectively
implemented by looking through all the plans produced and filtering
out the ones that do not meet the desired requirements. In the
future we intend to study algorithms that do such filtering online,
rather than after initial solutions are found.

@@
